\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-28 chapters/bttda.tex
Layout BTTDA

\chapter{Block-term tensor discriminant analysis}
\label{sec:bttda}
\input{chapters/bttda/tensorviz.tex}
\epigraph{%
  ``I journeyed to the market to purchase a cow, its price set at ninety-and-seven coins.
  Lacking this sum, I therefore borrowed fifty coins from each of my two good friends.
  With the hundred coins, I paid for the beast and received three coins in change.
  Of these, I returned one coin each to my companions, keeping one for myself.
  Now, I owe each friend forty-and-nine coins, and with mine one remaining, the
  total merely equals ninety-and-nine coins, short of one hundred!
  Tell me then, kind stranger, whither hath the missing coin gone?''
}{%
  Ancient Algerian riddle
}
\noindent\emph{This chapter was submitted as~\textcite{VanDenKerchove2024a}.}

@@

\subsection{\Acl{hoda}}
\Acl{hoda}~\cite{Phan2010} is a
supervised tensor-based feature extraction technique.
For a set of $N$ tensors of order $K$
$\left\{\ten{X}(n)\in\mathbb{R}^{D_1\times D_2 \times \cdots \times
		D_K}\right\}_n^N$, HODA finds projection matrices $\mat{U_k}$ for each mode $k$
that project a given $\ten{X}$ to a latent tensor
$\ten{G}\in\mathbb{R}^{r_1\times r_2\times\cdots\times r_K}$, usually with lower
dimensionality $(r_1\leq D_1,r_2\leq D_2,\ldots,r_K\leq D_K)$ using
tensor-matrix mode products:
\begin{equation}
	\ten{G}  = \ten{X}\mmpr{\mat{U}}
	\label{eq:HODA-backward}
\end{equation}
as visualized in \cref{fig:HODA-backward}.
\begin{figure}[t]
	\centering
	\input{figures/bttda/HODA_backward.tikz.tex}
  \caption[A \acs{hoda} backward projection.]{%
    A visualization of the multilinear projection obtained by \acf{hoda} applied to a third-order tensor
    sample $\ten{X}$ with shape $(D_1,D_2, D_3)$.
		\Ac{hoda} finds projection matrices $\mat{U}_k$ such that maximal
		discriminability between classes is achieved in the projected latent tensors
		$\ten{G}$ with reduced dimensionality $(r_1,r_2,r_3)$.}
	\label{fig:HODA-backward}
\end{figure}

@@

The \ac{hoda} projection in \cref{eq:HODA-backward} is an example
of a backward model.
A straightforward and computationally efficient candidate for a corresponding
forward model, visualized in \cref{fig:HODA-forward}, is given by:
\begin{equation}
  \ten{X} = \ten{G}\mmpr{\mat{A}^\intercal} + \ten{\mathbfcal{E}} =
  \hat{\ten{X}} + \ten{\mathbfcal{E}}
	\label{eq:HODA-forward}
\end{equation}
with \emph{activation patterns} $\mat{A}_k \in \mathbb{R}^{D_k\times r_k}$,
reconstructed tensor $\hat{\ten{X}}$, and error term $\ten{\mathbfcal{E}}$.
\begin{figure}[t]
	\centering
	\input{figures/bttda/HODA_forward.tikz.tex}
  \caption[A forward projection for \acs{hoda}.]{The forward projection for HODA.
    By calculating activation patterns $\mat{A}_k$, the original tensor $\ten{X}$ can approximately be
    reconstructed as $\hat{\ten{X}}$ from projected latent tensor $\ten{G}$.
    The reconstruction is accurate up to an error term $\ten{\mathbfcal{E}}$,
    such that $\ten{X}=\hat{\ten{X}}+\ten{\mathbfcal{E}}$.
    $\mat{A}_k$ are chosen such that the variability captured in the latent tensor is
    maximally expressed in the reconstructed tensor $\hat{\ten{X}}$ and not in
    the error term $\ten{\mathbfcal{E}}$.}
	\label{fig:HODA-forward}
\end{figure}

@@

We thus extend the \ac{hoda} feature extraction scheme to \acf{bttda}.
\Ac{bttda} finds multiple discriminative blocks, such that its forward
model adheres to the block-term tensor structure:
\begin{align}
	\ten{X} & = \sum_b^B\ten{G}^{(b)}\mmpr{\mat{A}^{(b)}} + \ten{\mathbfcal{E}}
	\label{eq:BTTDA-forward}
\end{align}
for $B$ extracted latent tensors $\ten{G}^{(b)}$ and residual error term
$\ten{\mathbfcal{E}}$.
The BTTDA model is further illustrated by~\cref{fig:BTTDA}.
\begin{figure*}[t]
	\centering
	\input{figures/bttda/BTTDA.tikz.tex}
  \caption[A forward model for \acs{bttda}.]{A forward model for \acf{bttda}.
    \Ac{bttda} can extract more features
    than \ac{hoda} by iteratively finding a latent tensor $\ten{G}^{(b)}$ in a
		deflation scheme.
    The \ac{hoda} backward projection is first applied. Next, the
		input data is reconstructed via the HODA forward model and the
		difference between the two is found.
		Finally, this process is repeated with this difference as input data, until a
		desired number of blocks $B$ has been found.}
	\label{fig:BTTDA}
\end{figure*}
The block-term structure of this model implies that it is a generalization of both
the Tucker-structured \ac{hoda} and PARAFAC-structured discriminant feature
extraction.
If $B$ in \cref{eq:BTTDA-forward} is set to one, \ac{bttda} is equivalent to
\ac{hoda}; if at each term $b$ the rank of the core tensor
$(r_1^{(b)}=r_2^{(b)}=\ldots=r_k^{(b)}=1)$, a PARAFAC structure is assumed.

@@

\section{Experiments}
\subsection{Datasets and decoders}
We evaluated our proposed model in two offline EEG-based BCI decoding problems:
the event-related potential (ERP) and motor imagery (MI) paradigms using a
selection of the openly available MOABB benchmarking datasets~\cite{Aristimunha2023}.
Two ERP and two MI datasets were retained to reduce computational
demand.
Details about these datasets are found in \cref{tab:moabb}.
For the ERP datasets, the task is to distinguish target from non-target ERPs,
while the MI datasets consist of distinguishing different imagined limb
movements.
Within-session classification performance was assessed using stratified 5-fold
cross-validation to calculate the area under the receiver operating
characteristic curve (ROC-AUC).
\begin{table*}[t]
	\centering
	\footnotesize
  \makebox[\textwidth][c]{
	  \input{chapters/bttda/moabb.tex}
  }
  \caption[MOABB datasets used for evaluation.]{MOABB datasets used for evaluation, with the number of
		subjects (\# Sub.), the number of EEG channels (\# Chan.), the number of
    classes (\# Classes), the number of trials or trials per class for ERP
    datasets (\# Trials), the epoch length (Epoch len.), the sampling
		frequency (S. freq.), the number of sessions per subject (\# Sess.) and the
		number of runs (\# Runs).
    Adapted from \textcite{Aristimunha2023} and \textcite{Chevallier2024}.}
	\label{tab:moabb}
\end{table*}

@@

\Cref{tab:erp-score} lists the cross-validated ROC-AUC for all evaluated
decoders.
\begin{table}[t]
	\footnotesize
	\centering
	\input{chapters/bttda/score_erp.tex}
  \caption[Within-session classification score for 2 event-related potential
  datasets.]{Area under the receiver operating characteristic curve for
		cross-validated within-session evaluation for HODA and our proposed decoders
		PARAFACDA and BTTDA evaluated on 2 event-related potential datasets.
    Scores for other decoders were taken from \textcite{Chevallier2024}.
		BTTDA reaches the highest performance for both evaluated datasets, closely
		followed by PARAFACDA.
	}
	\label{tab:erp-score}
\end{table}
The highest performance is achieved with the proposed BTTDA+LDA.
One-sided Wilcoxon signed-rank tests with a significance level of $\alpha=0.05$
reveal that both PARAFACDA+LDA and BTTDA+LDA significantly outperform
HODA+LDA in both the BNCI2014-008 (PARAFACDA+LDA: $p=0.0039$, BTTDA+LDA:
$p=0.0039$) and the BNCI2015-003 (PARAFACDA+LDA: $p=0.0001$, BTTDA+LDA:
$p=0.0049$) datasets.
No significant difference was found between BTTDA+LDA and PARAFACDA+LDA.

@@

\cref{tab:mi-score} lists the cross-validated classification scores for
the evaluated motor imagery datasets.
Note that, in line with~\textcite{Chevallier2024}, accuracy is reported for the
multi-class classification problem in BNCI2014-001, while ROC-AUC was reported for the
binary classification problem in BNCI2014-004.
\begin{table}[t]
  \centering
	\footnotesize
	\input{chapters/bttda/score_mi.tex}
  \caption[Within-session classification score for 2 motor imagery datasets.]{Classification score for
		cross-validated within-session evaluation for HODA+LDA and our proposed decoders
		PARAFACDA+LDA and BTTDA+LDA evaluated on 2 motor imagery datasets.
		Scores for other decoders were taken from \textcite{Chevallier2024}.
    Accuracy is listed for BNCI2014-001, and \ac{rocauc} for BNCI2014-004.
    BTTDA outperforms HODA and PARAFACDA for BNCI2014-001 but does not reach
    performance comparable to the current state-of-the-art.
    For BNCI2014-004, PARAFACDA and BTTDA perform
    approximately on par with HODA and the current state-of-the art.}
	\label{tab:mi-score}
\end{table}
For the BNCI2014-001 dataset, HODA+LDA and our proposed decoders do not reach
satisfactory performance compared to the comparison methods, yet both
PARAFACDA+LDA and BTTDA+LDA improve upon HODA+LDA, and for BTTDA+LDA this
difference is significant ($p=0.0269$).
For the BNCI2014-004 dataset, the performance gap with comparison methods is
smaller, but no significant differences were found between HODA+LDA,
PARAFACDA+LDA, and BTTDA+LDA.

\subsection{Block contribution}
To analyze the contribution of extra feature blocks extracted by BTTDA over
the first one found by HODA, we pick an ERP ($K=2$) dataset
(BNCI2014-008) and an MI ($K=3$) dataset (BNCI2014-001).
We report within-session ROC-AUC scores for training, validation, and test data as a function
of the number of blocks, shown in \cref{fig:blocks}.
Training and validation folds were taken from the model selection procedure.
Additionally, the Normalized Mean Squared Error (NMSE) is reported for the
reconstructed from the truncated BTTDA decomposition
$\textstyle{\hat{\ten{X}}^{(B)}=\sum_b^B\ten{G}^{(b)}\mmpr{\mat{U}^{(b)}}}$.
NMSE is calculated as:
\begin{equation}
  \nmse\left(\ten{X}, \hat{\ten{X}}^{(B)}\right) =
	\frac{\sum_n^N\left\lVert\ten{X}(n)-\ten{X}_\text{rec}^{(B)}(n)\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{X}(n)\right\rVert_F^2}
\end{equation}
\begin{figure*}[t]
  \centering
  \makebox[\linewidth][c]{
	  \input{figures/bttda/blocks.pgf}
  }
  \caption[Analysis of NMSE and classification score per block.]{%
    Normalized Mean Square Error (NMSE) (left), and difference in
    classification score for the training and validation for the ERP datasets
    (top row) and the MI dataset (bottom row)	of the greedy model selection
    procedure (right), as a function of the number of BTTDA blocks $b$.
		While NMSE monotonically decreases for the evaluated datasets, better class
		separation will be achieved, but eventually overfitting occurs and validation
		and test scores will drop, or plateau due to feature selection.
	}
	\label{fig:blocks}
\end{figure*}

@@

Additionally, the forward modeling step inherent to BTTDA results
in an interpretable model since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
\Cref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-0.png}
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-1.png}
  \caption[Extracted \acs{bttda} activation patterns.]{%
    Spatial (left two columns) and temporal (middle column) activation patterns and
		condition contrasts (right column) obtained after forward projection of the latent
    features for 2 blocks (row 1 and row 2) of rank $(2,2)$ of \ac{bttda}
    fit on the full dataset BNCI2014-008.
    The separate blocks approximately model different \ac{erp}
		components.}
	\label{fig:forward}
\end{figure*}
Given informed or correctly tuned hyperparameters, this method could be used to
e.g., separate \ac{erp} components or neural processes based on the task-related
information in the class labels.

>>>>

\chapter{Block-term tensor discriminant analysis}
\label{sec:bttda}
\input{chapters/bttda/tensorviz.tex}
\epigraph{%
  ``I journeyed to the market to purchase a cow, its price set at ninety-and-seven coins.
  Lacking this sum, I therefore borrowed fifty coins from each of two good friends.
  With these hundred coins, I paid for the beast and received three coins in change.
  Of the three, I returned one coin each to my companions, keeping one for myself.
  Now, I owe each friend forty-and-nine coins, a total debt of
  ninety-and-eight. And with mine one remaining, the
  total merely equals ninety-and-nine coins, short of the hundred I borrowed!
  Tell me then, kind stranger, whither hath the missing coin gone?''
}{%
  Ancient Algerian riddle
}
\publicationnote{This chapter was submitted as~\textcite{VanDenKerchove2024a}.}

@@

\subsection{\Acl{hoda}}
\Acl{hoda}~\cite{Phan2010} is a
supervised tensor-based feature extraction technique.
For a set of $N$ tensors of order $K$
$\left\{\ten{X}(n)\in\mathbb{R}^{D_1\times D_2 \times \cdots \times
		D_K}\right\}_n^N$, HODA finds projection matrices $\mat{U_k}$ for each mode $k$
that project a given $\ten{X}$ to a latent tensor
$\ten{G}\in\mathbb{R}^{r_1\times r_2\times\cdots\times r_K}$, usually with lower
dimensionality $(r_1\leq D_1,r_2\leq D_2,\ldots,r_K\leq D_K)$ using
tensor-matrix mode products:
\begin{equation}
	\ten{G}  = \ten{X}\mmpr{\mat{U}}
	\label{eq:HODA-backward}
\end{equation}
as visualized in \cref{fig:HODA-backward}.
\begin{figure}[t]
	\input{figures/bttda/HODA_backward.tikz.tex}
  \caption[A \acs{hoda} backward projection.]{%
    A visualization of the multilinear projection obtained by \acf{hoda} applied to a third-order tensor
    sample $\ten{X}$ with shape $(D_1,D_2, D_3)$.
		\Ac{hoda} finds projection matrices $\mat{U}_k$ such that maximal
		discriminability between classes is achieved in the projected latent tensors
		$\ten{G}$ with reduced dimensionality $(r_1,r_2,r_3)$.}
	\label{fig:HODA-backward}
\end{figure}

@@

The \ac{hoda} projection in \cref{eq:HODA-backward} is an example
of a backward model.
A straightforward and computationally efficient candidate for a corresponding
forward model, visualized in \cref{fig:HODA-forward}, is given by:
\begin{equation}
  \ten{X} = \ten{G}\mmpr{\mat{A}^\intercal} + \ten{\mathbfcal{E}} =
  \hat{\ten{X}} + \ten{\mathbfcal{E}}
	\label{eq:HODA-forward}
\end{equation}
with \emph{activation patterns} $\mat{A}_k \in \mathbb{R}^{D_k\times r_k}$,
reconstructed tensor $\hat{\ten{X}}$, and error term $\ten{\mathbfcal{E}}$.
\begin{figure}[t]
	\input{figures/bttda/HODA_forward.tikz.tex}
  \caption[A forward projection for \acs{hoda}.]{The forward projection for HODA.
    By calculating activation patterns $\mat{A}_k$, the original tensor $\ten{X}$ can approximately be
    reconstructed as $\hat{\ten{X}}$ from projected latent tensor $\ten{G}$.
    The reconstruction is accurate up to an error term $\ten{\mathbfcal{E}}$,
    such that $\ten{X}=\hat{\ten{X}}+\ten{\mathbfcal{E}}$.
    $\mat{A}_k$ are chosen such that the variability captured in the latent tensor is
    maximally expressed in the reconstructed tensor $\hat{\ten{X}}$ and not in
    the error term $\ten{\mathbfcal{E}}$.}
	\label{fig:HODA-forward}
\end{figure}

@@

We thus extend the \ac{hoda} feature extraction scheme to \acf{bttda}.
\Ac{bttda} finds multiple discriminative blocks, such that its forward
model adheres to the block-term tensor structure:
\begin{align}
	\ten{X} & = \sum_b^B\ten{G}^{(b)}\mmpr{\mat{A}^{(b)}} + \ten{\mathbfcal{E}}
	\label{eq:BTTDA-forward}
\end{align}
for $B$ extracted latent tensors $\ten{G}^{(b)}$ and residual error term
$\ten{\mathbfcal{E}}$.
The BTTDA model is further illustrated by~\cref{fig:BTTDA}.
\begin{figure*}[t]
	\input{figures/bttda/BTTDA.tikz.tex}
  \caption[A forward model for \acs{bttda}.]{A forward model for \acf{bttda}.
    \Ac{bttda} can extract more features
    than \ac{hoda} by iteratively finding a latent tensor $\ten{G}^{(b)}$ in a
		deflation scheme.
    The \ac{hoda} backward projection is first applied. Next, the
		input data is reconstructed via the HODA forward model and the
		difference between the two is found.
		Finally, this process is repeated with this difference as input data, until a
		desired number of blocks $B$ has been found.}
	\label{fig:BTTDA}
\end{figure*}
The block-term structure of this model implies that it is a generalization of both
the Tucker-structured \ac{hoda} and PARAFAC-structured discriminant feature
extraction.
If $B$ in \cref{eq:BTTDA-forward} is set to one, \ac{bttda} is equivalent to
\ac{hoda}; if at each term $b$ the rank of the core tensor
$(r_1^{(b)}=r_2^{(b)}=\ldots=r_k^{(b)}=1)$, a PARAFAC structure is assumed.

@@

\section{Experiments}
\subsection{Datasets and decoders}
We evaluated our proposed model in two offline EEG-based BCI decoding problems:
the event-related potential (ERP) and motor imagery (MI) paradigms using a
selection of the openly available MOABB benchmarking datasets~\cite{Aristimunha2023}.
Two ERP and two MI datasets were retained to reduce computational
demand.
Details about these datasets are found in \cref{tab:moabb}.
For the ERP datasets, the task is to distinguish target from non-target ERPs,
while the MI datasets consist of distinguishing different imagined limb
movements.
Within-session classification performance was assessed using stratified 5-fold
cross-validation to calculate the area under the receiver operating
characteristic curve (ROC-AUC).
\begin{table*}[t]
  \makebox[\textwidth][c]{
	  \input{chapters/bttda/moabb.tex}
  }
  \caption[MOABB datasets used for evaluation.]{MOABB datasets used for evaluation, with the number of
		subjects (Sub.), the number of EEG channels (Chan.), the number of
    classes (Cls.), the number of trials or trials per class for ERP
    datasets (Trials), the epoch length (Epoch len.), the sampling
		frequency (S. freq.), the number of sessions per subject (Sess.) and the
		number of runs.
    Adapted from \textcite{Aristimunha2023} and \textcite{Chevallier2024}.}
	\label{tab:moabb}
\end{table*}

@@

\Cref{tab:erp-score} lists the cross-validated ROC-AUC for all evaluated
decoders.
\begin{table}[t]
	\input{chapters/bttda/score_erp.tex}
  \caption[Within-session classification score for 2 event-related potential
  datasets.]{Area under the receiver operating characteristic curve for
		cross-validated within-session evaluation for HODA and our proposed decoders
		PARAFACDA and BTTDA evaluated on 2 event-related potential datasets.
    Scores for other decoders were taken from \textcite{Chevallier2024}.
		BTTDA reaches the highest performance for both evaluated datasets, closely
		followed by PARAFACDA.
	}
	\label{tab:erp-score}
\end{table}
The highest performance is achieved with the proposed BTTDA+LDA.
One-sided Wilcoxon signed-rank tests with a significance level of $\alpha=0.05$
reveal that both PARAFACDA+LDA and BTTDA+LDA significantly outperform
HODA+LDA in both the BNCI2014-008 (PARAFACDA+LDA: $p=0.0039$, BTTDA+LDA:
$p=0.0039$) and the BNCI2015-003 (PARAFACDA+LDA: $p=0.0001$, BTTDA+LDA:
$p=0.0049$) datasets.
No significant difference was found between BTTDA+LDA and PARAFACDA+LDA.

@@

\cref{tab:mi-score} lists the cross-validated classification scores for
the evaluated motor imagery datasets.
Note that, in line with~\textcite{Chevallier2024}, accuracy is reported for the
multi-class classification problem in BNCI2014-001, while ROC-AUC was reported for the
binary classification problem in BNCI2014-004.
\begin{table}[t]
	\input{chapters/bttda/score_mi.tex}
  \caption[Within-session classification score for 2 motor imagery datasets.]{Classification score for
		cross-validated within-session evaluation for HODA+LDA and our proposed decoders
		PARAFACDA+LDA and BTTDA+LDA evaluated on 2 motor imagery datasets.
		Scores for other decoders were taken from \textcite{Chevallier2024}.
    Accuracy is listed for BNCI2014-001, and \ac{rocauc} for BNCI2014-004.
    BTTDA outperforms HODA and PARAFACDA for BNCI2014-001 but does not reach
    performance comparable to the current state-of-the-art.
    For BNCI2014-004, PARAFACDA and BTTDA perform
    approximately on par with HODA and the current state-of-the art.}
	\label{tab:mi-score}
\end{table}
For the BNCI2014-001 dataset, HODA+LDA and our proposed decoders do not reach
satisfactory performance compared to the comparison methods, yet both
PARAFACDA+LDA and BTTDA+LDA improve upon HODA+LDA, and for BTTDA+LDA this
difference is significant ($p=0.0269$).
For the BNCI2014-004 dataset, the performance gap with comparison methods is
smaller, but no significant differences were found between HODA+LDA,
PARAFACDA+LDA, and BTTDA+LDA.

\subsection{Block contribution}
To analyze the contribution of extra feature blocks extracted by BTTDA over
the first one found by HODA, we pick an ERP ($K=2$) dataset
(BNCI2014-008) and an MI ($K=3$) dataset (BNCI2014-001).
We report within-session ROC-AUC scores for training, validation, and test data as a function
of the number of blocks, shown in \cref{fig:blocks}.
Training and validation folds were taken from the model selection procedure.
Additionally, the Normalized Mean Squared Error (NMSE) is reported for the
reconstructed from the truncated BTTDA decomposition
$\textstyle{\hat{\ten{X}}^{(B)}=\sum_b^B\ten{G}^{(b)}\mmpr{\mat{U}^{(b)}}}$.
NMSE is calculated as:
\begin{equation}
  \nmse\left(\ten{X}, \hat{\ten{X}}^{(B)}\right) =
	\frac{\sum_n^N\left\lVert\ten{X}(n)-\ten{X}_\text{rec}^{(B)}(n)\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{X}(n)\right\rVert_F^2}
\end{equation}
\begin{figure*}[t]
  \sffamily
	\input{figures/bttda/blocks.pgf}
  \caption[Analysis of NMSE and classification score per block.]{%
    Normalized Mean Square Error (NMSE) (left), and difference in
    classification score for the training and validation for the ERP datasets
    (top row) and the MI dataset (bottom row)	of the greedy model selection
    procedure (right), as a function of the number of BTTDA blocks $b$.
		While NMSE monotonically decreases for the evaluated datasets, better class
		separation will be achieved, but eventually overfitting occurs and validation
		and test scores will drop, or plateau due to feature selection.
	}
	\label{fig:blocks}
\end{figure*}

@@

Additionally, the forward modeling step inherent to BTTDA results
in an interpretable model since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
\Cref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-0_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-0_contrast.png}
    \end{minipage}
    \vspace{-1em}
    \subcaption{The first block.}
  \end{subfigure}

  \begin{subfigure}[b]{\linewidth}
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-0.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.15\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-sp-1.pgf}
      \vspace{3em}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.25\linewidth}
      \inputpgf{figures/bttda}{forward_block-1_ap-tmp.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.3\linewidth}
      \includegraphics[width=\linewidth]{figures/bttda/forward_block-1_contrast.png}
    \end{minipage}
    \vspace{-1em}
    \subcaption{The second block.}
  \end{subfigure}

	%\includegraphics[width=\linewidth]{figures/bttda/forward_block-0.png}
	%\includegraphics[width=\linewidth]{figures/bttda/forward_block-1.png}
  \caption[Extracted \acs{bttda} activation patterns.]{%
    Spatial (left two columns) and temporal (middle column) activation patterns and
		condition contrasts (right column) obtained after forward projection of the latent
    features for 2 blocks of rank $(2,2)$ of \ac{bttda}
    fit on the full dataset BNCI2014-008.
    The separate blocks approximately model different \ac{erp}
		components.}
	\label{fig:forward}
\end{figure*}
Given informed or correctly tuned hyperparameters, this method could be used to
e.g., separate \ac{erp} components or neural processes based on the task-related
information in the class labels.
